---
title: "Development actions history"
output: html_document
editor_options: 
  chunk_output_type: console
---

All commands that you use to use when developing packages...

# First time just after creating the project

- Fill the following chunk to create the DESCRIPTION of your package

```{r description}
# Describe your package
fusen::fill_description(
  pkg = here::here(),
  fields = list(
    Title = "Tools for simulation analyses",
    Description = "Includes a suite of tools to simulate virtual ecological systems, from environment, to animals and surveys.",
    `Authors@R` = c(
      person("Charlotte", "Lambert", email = "charlotte.lambert@univ-lr.fr", role = c("aut", "cre"), comment = c(ORCID = "0000-0002-1128-5262"))
    )
  )
)
# Define License with use_*_license()
usethis::use_mit_license("Charlotte Lambert")
```

# All-in-one function to deploy publicly on GitHub

Either use this GitHub all-in-one function or choose the steps one-by-one in the following sections.
See `vignette("share-on-a-github-website", package = "fusen")` for more information.

```{r, eval=FALSE}
# _All-in-one share of your package and its documentation on GitHub
fusen::init_share_on_github()
```


# Start using git

```{r, eval=FALSE}
## definit le compte
usethis::use_git_config(scope = "user",
                        user.name = "CLambert1",
                        user.email = "charlotte.anne.lambert@gmail.com",
                        credential.helper = "store",
                        init.defaultBranch = "main")
# donne les identifiants
gitcreds::gitcreds_set(url = "https://github.com/")

# lance l'utilisation de git en local
usethis::use_git()

# associe au repo en remote
# usethis::use_git_remote(url = "https://github.com/CLambert1/simtools.git") # ne fonctionne pas car l'autentification via rstudio n'est plus autorisé par github
# doit faire dans le terminal : git clone https://github.com/CLambert1/simtools.git ; puis autentification dans le pop-up
# ensuite git push -u origin main dans le terminal pour faire le premier push et activer la syncro
  
# Deal with classical files to ignore
usethis::git_vaccinate()
# Use main for primary branch
usethis::git_default_branch_rename()
```

# Set extra sources of documentation

```{r, eval=FALSE}
# Install a first time
remotes::install_local()
# README
usethis::use_readme_rmd()
# Code of Conduct
usethis::use_code_of_conduct("charlotte.lambert@univ-lr.fr")
# NEWS
usethis::use_news_md()
```

**From now, you will need to "inflate" your package at least once to be able to use the following commands. Let's go to your flat template, and come back here later if/when needed.**

# Mettre des données dans mon package

```{r, eval=FALSE}
# add data-raw/ dans le dossier du package - use once
usethis::use_data_raw() # creer le dossier et DATASET.R ; ensuite créer des scripts R pour les transformer

```


# Package development tools
## Use once

```{r, eval=FALSE}
# Pipe
usethis::use_pipe()

# package-level documentation
usethis::use_package_doc()

# update dependancies etc
attachment::att_amend_desc()

# create a vignette for the whole package
usethis::use_vignette("simtools")


# GitHub
# Add your credentials for GitHub
usethis::create_github_token()
gitcreds::gitcreds_set()

# _Or one step at a time
# Send your project to a new GitHub project (included in `init_share_on_github()`)
usethis::use_github()

# Set Continuous Integration
# _GitHub (included in `init_share_on_github()`)
usethis::use_github_action_check_standard()
usethis::use_github_action("pkgdown") # déployement en ligne du site sur une Github page via GutHub Action (publie dans une branche 'gh_pages')
usethis::use_github_action("test-coverage")
# _GitLab
gitlabr::use_gitlab_ci(type = "check-coverage-pkgdown")
```

## Use everytime needed

```{r, eval=FALSE}
# Simulate package installation
pkgload::load_all()

# Generate documentation and deal with dependencies
attachment::att_amend_desc()

# Check the package
devtools::check()

# Add a new flat template
fusen::add_flat_template("add")

# monter version - choisi les niveaux mais doit maintenir NEWS.md à la main pour noter les nouveautés. 
# peut aussi bien faire directement à la main dans NEWS et dans description, c'est aussi simple. 
usethis::use_version()

# tester le taux de couverture du package
covr::package_coverage()
# quelles parties ne sont pas couvertes
my_coverage <- covr::package_coverage()
covr::zero_coverage(my_coverage)

# pour créer une page exprès au sujet du taux de couverture des tests unitaires
covr::report()
```

# Share the package

```{r, eval=FALSE}
# set and try pkgdown documentation website locally
usethis::use_pkgdown()

# reconstruit le site en entier
pkgdown::build_site(override = list(destination = "inst/site")) # faire dans inst/ plutot que à la racine, pour que le site soit installé en local avec le package et donc accessible sans internet
# custom appearence in _pkgdown.yml (racine du package)
# choose bootswatch name from : https://bootswatch.com/
# see https://pkgdown.r-lib.org/articles/customise.html for customisation options
# re-build or update the site (needs refreshing the browser to see it)
pkgdown::init_site()
pkgdown::build_home_index()


# inflate all flat files all at once (must have been inflated separately at least once)
fusen::inflate_all()

# update readme for its inclusion in the website
devtools::build_readme()

# build the tar.gz with vignettes to share with others
devtools::build(vignettes = TRUE)

# Share your package and its documentation on GitHub
# usethis::create_github_token()
# gitcreds::gitcreds_set()
fusen::init_share_on_github()
```

# check l'installation du package

```{r}
devtools::build(vignettes = TRUE) # construit le package au format tar.gz
remotes::install_local(path = "~/simtools_0.0.0.9000.tar.gz") # installe le package depuis le format tar.gz
simtools::launch_help()
```

