---
title: "flat_simulate_mvmt.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

```{r function-utils}
#' Return random direction angles
#'
#' @inheritParams circular::rwrappednormal 
#'
#' @importFrom circular rwrappednormal 
#'
#' @return n random angles
#' @export
#'
randomdir <- function(n,mu,kappa){
  out <- circular::rwrappednormal(10, mu = circular::circular(mu, units = "degrees", 
                                                              template = "geographic"),
                                 sd = kappa, control.circular = list()) %% 360
  return(as.double(out))
}

#' Return random step lengths
#'
#' @inheritParams stats::rgamma 
#'
#' @importFrom stats rgamma 
#'
#' @return n random step lengths
#' @export
#'
randomdist <- function(n, shape, rate){
  out <- stats::rgamma(n, shape, rate)
  return(out)
}
```

# Define potential next positions

\code{\link{potential_position_func}} generates a bunch of potential positions to chose the next step from. Positions are generated based on provided bearing (Von Mises distribution) and step distributions (gamma distribution). For each positions, the value of env_rast is retrieved as well as the distance from the colony location. 

```{r function-potential_position_func}
#' Identify potential next positions
#'
#' @param n Numeric, the number of potential positions to generate
#' @param bearing Numeric vector, mu and kappa to be passed on randomdir
#' @param step  Numeric, shape and rate to be passed on randomdist
#' @param from data.frame including Lon and Lat fields and stepID of the point of origin
#' @param colony_location data.frame vector with lon and lat position of the colony
#' @param resource.layer SpatRaster, the suitability or any resource layer to be used to sort potential position (a single layer)
#' 
#' @importFrom terra extract nlyr
#' @importFrom fields rdist
#' 
#' @return A data.frame with the n potential positions from the point of origin, informed with the value of the provided environmental layer corresponding to each as well as the distance from the colony. 
#' @export
#'
#' @examples
potential_position_func <- function(n, bearing, step, from, colony_location, resource_layer){
  if(isTRUE(inherits(resource_layer, "SpatRaster"))){
    if(terra::nlyr(resource_layer) > 1){
      stop("env_layer must include a single layer")
    }
  } else { stop("env_layer must be a SpatRaster") }
  if(isFALSE(inherits(from, "data.frame"))){
    stop("from must be a data.frame")
  } else {
    if(isFALSE(nrow(from) == 1)){
      stop(glue::glue("from must include a single row"))
    } else {
      if(isFALSE(all(c("Lon", "Lat") %in% names(from)))){
      stop(glue::glue("from must include Lon, Lat columns"))
    }}
  }
  if(length(bearing) != 2){
    stop("bearing must be of length 2")
  }
  if(length(step) != 2){
    stop("step must be of length 2")
  }
  df <- data.frame(angle = randomdir(n = n, mu = bearing[1], kappa = bearing[2]),
                   step = randomdist(n = n, shape = step[1], rate = step[2]))      # generate angles in degree
  df$Lon <- from$Lon + (cos(df$angle) * df$step)
  df$Lat <- from$Lat + (sin(df$angle) * df$step)
  df$stepID <- from$stepID
  df$env <- terra::extract(resource_layer, df[, c("Lon","Lat")])[[2]]
  df$dist_col <- as.vector(fields::rdist(colony_location, df[, c("Lon", "Lat")]))
  df$potential_position_id <- 1:n
  return(df)
}

```

```{r example-potential_position_func}
from <- data.frame(Lon = 1, Lat = 2, toto = 5)
colony_location <- data.frame(Lon = 2, Lat = 5)
cdt <- generate_env_layer(grid = create_grid(), n = 1)$rasters
potential_position_func(n = 10, from = from, 
                        colony_location = colony_location, resource_layer = cdt, 
                        bearing = c(90,10), step = c(4.5, 3))
```
  
```{r tests-potential_position_func}
test_that("potential_position_func works", {
  expect_true(inherits(potential_position_func, "function")) 
  
  from <- data.frame(Lon = 1, Lat = 2, toto = 5)
  colony_location <- data.frame(Lon = 2, Lat = 5)
  cdt <- generate_env_layer(grid = create_grid(), n = 1)$rasters

  # check returns an error if env_rast is not a raster
  expect_error(
    object = potential_position_func(from = from, 
                                     colony_location = colony_location, resource_layer = from)
  )
  # check returns an error if from does not include coord_col names
  expect_error(
    object = potential_position_func(from = from[,c("Lon","toto")], 
                                     colony_location = colony_location, resource_layer = cdt)
  )
  # check returns an error when bearing is not of length 2
  expect_error(
    object = potential_position_func(from = from, coord_col = coord_col,
                                     colony_location = colony_location, resource_layer = cdt,
                                     bearing = 1, step = c(2,3))
  )
})
```


# Simulate the trajectory of a single central-place forager

\code{\link{simulate_trajectory_CPF}} simulates a full foraging trajectory of an individual departing from its colony. The individual launches its trip at a given time period (starting.hour, ymd_hms format) and starts its returning trip either when at a given distance from its colony (provided with max_dist) or when an hour before sunset. The user can define the bearing and step distributions for the departing step, but also for the travelling and foraging bouts. By default, travelling are set to be directed movement patterns (large steps with low variability, low angles with low variability) and foraging to be area-restricted movements (short steps with low variability, large angles with large variability). 
The user can define the step duration, that is the time interval at which the positions are sampled.  

For each position, ten potential subsequent positions are randomly sampled using \code{\link{potential_position_func()}}, that is with movement parameters based on the activity the individual is engaged in at the previous step (travelling or foraging). The environmental suitability (a raster layer provided with resource_layer) is retrieved for each potential point based on their geographic coordinates, and their distance from the colony is computed using the rdist function from the fields package. 

The first four positions of the track after colony departure are randomly selected within the ten potential points sampled as described above, using travelling movement. Afterwards, subsequent positions are randomly sampled among the four out of ten potential points which have the highest values of environmental suitability. If the environmental suitability exceeds a given threshold (activity_threshold), the individual switches to foraging movements; if not, the individual continues travelling.

When individuals reaches the threshold distance to colony (max_dist), or the trip duration exceeds the given duration (max_duration), they start their homeward journey. Potential next positions are sampled using only travelling movement parameters, and the next positions are selected based on the minimum distance to the colony. If several potential positions meet the selection criteria, the next step is randomly sampled within these positions. An individual is considered back when within 0.5 spatial unit (the unit depends on the provided resource layer) from the colony.

```{r function-simulate_trajectory_CPF}
#' Simulate individual trajectory
#'
#' @param initial_position data.frame containing the coordinates of the point the individual starts from (Lon, Lat fields), considered to be the colony location
#' @param resource_layer SpatRaster of the resource layer
#' @param starting_hour Starting hour
#' @param starting_bearing Bearing when departing the colony (mu, kappa)
#' @param starting_step Step length when departing the colony (mu, kappa)
#' @param travel_bearing Bearing from previous step when traveling (mu, kappa)
#' @param travel_step Distance from previous step when traveling (scale, rate) 
#' @param foraging_bearing Bearing from previous step when foraging/searching (mu, kappa)
#' @param foraging_step Distance from previous step when foraging (scale, rate)
#' @param minx Minimum longitude of the bounding box to simulate trajectory within
#' @param maxx Maximum longitude of the bounding box to simulate trajectory within
#' @param miny Minimum latitude of the bounding box to simulate trajectory within
#' @param maxy Maximum latitude of the bounding box to simulate trajectory within
#' @param max_dist Distance threshold value above which the individual starts its homing bout
#' @param max_duration Total duration of a full trip, in minutes (defaults to 720, 12h)
#' @param step_duration Duration of a step length, in minutes (defaults 1), defines the sampling interval
#' @param activity_threshold Threshold environmental value defining when an individual switches from travelling to foraging movement
#' 
#'
#' @import lubridate
#' @importFrom dplyr full_join slice_sample slice_head desc arrange
#' 
#' @return a data.frame
#' @export
#'
#' @examples
simulate_trajectory_CPF <- function(initial_position, 
                                resource_layer, 
                       starting_hour,
                       starting_bearing = c(90,10),
                       starting_step = c(4.5, 3),
                       travel_bearing = c(0, 20), # (mu, kappa) -> higher kappa = higher concentration
                       travel_step = c(3, 3), #  (scale, rate) -> smaller the rate, higher the dispersion
                       foraging_bearing = c(0, 0.5), 
                       foraging_step = c(1, 3), 
                       minx = 0, maxx = 90, 
                       miny = 0, maxy = 90,
                       max_dist = 40,
                       step_duration = 1,
                       activity_threshold = 0.5,
                       max_duration = 720) 
{
  if(isFALSE(initial_position$Lon >= minx & initial_position$Lon <= maxx &
         initial_position$Lat >= miny & initial_position$Lat <= maxy)){
    stop("initial_position must be included in the bounding box given by minx, maxx, miny and maxy")
  }
  ext_coord <- as.vector(terra::ext(resource_layer))
  if(isFALSE(initial_position$Lon >= ext_coord[[1]] & initial_position$Lon <= ext_coord[[2]] &
            initial_position$Lat >= ext_coord[[3]] & initial_position$Lat <= ext_coord[[4]])){
    stop("initial_position must be included in the resource_layer extent")
  }
  if(nrow(initial_position) != 1){
    stop("initial_position must include a single row")
  }
  # check sur structure de initial_position sont fait dans potential_position_func()
  
  track_output <- data.frame(Lon = initial_position$Lon,
                             Lat = initial_position$Lat,
                             angle = NA,
                             step = 0,
                             env = NA,
                             dist_col = 0,
                             activity = "start",
                             stepID = 0,
                             ymd_hms = starting_hour,
                             x_from = initial_position$Lon,
                             y_from = initial_position$Lat,
                             date_from = starting_hour)
  
  ### Departs from the colony - outbound phase of travel and foraging bouts
  i <- 2
  while(
      track_output[i-1, "dist_col"] < max_dist &
      track_output[i-1, "ymd_hms"] < ymd_hms(starting_hour) + lubridate::minutes(max_duration)
  ){
    # starts with the initial bearing
    if(i == 2){ 
      potential_position <- potential_position_func(n = 10, 
                                                    bearing = starting_bearing, 
                                                    step = starting_step, 
                                                    from = track_output[i - 1, ],
                                                    colony_location = initial_position, 
                                                    resource_layer = resource_layer) |>
        subset(Lon < maxx & Lon > minx) |>
        subset(Lat < maxy & Lat > miny) |>
        dplyr::slice_sample(n = 1)
      if(is.null(nrow(potential_position))) { # if no points possible then do it again
        potential_position <- potential_position_func(n = 10, 
                                                      bearing = starting_bearing, 
                                                      step = starting_step, 
                                                      from = track_output[i - 1, ],
                                                      colony_location = initial_position, 
                                                      resource_layer = resource_layer) |>
          subset(Lon < maxx & Lon > minx) |>
          subset(Lat < maxy & Lat > miny) |>
          dplyr::slice_sample(n = 1)
        } 
      track_output[i, c("Lon","Lat","angle","step","env","dist_col")] <- potential_position[, c("Lon","Lat","angle","step","env","dist_col")]
      track_output[i, "stepID"] <-  track_output[i-1, "stepID"] + 1
      track_output[i, "ymd_hms"] <-  track_output[i-1, "ymd_hms"] + lubridate::minutes(step_duration)
      track_output[i, "x_from"] <-  track_output[i-1, "Lon"]
      track_output[i, "y_from"] <-  track_output[i-1, "Lat"]
      track_output[i, "date_from"] <-  track_output[i-1, "ymd_hms"]
      track_output[i, "activity"] <- "travel"
    }
    # continue with traveling parameters but not yet condition on environment
    if(i %in% c(3:5)){
      potential_position <- potential_position_func(n = 10, 
                                                    bearing = travel_bearing, 
                                                    step = travel_step, 
                                                    from = track_output[i - 1, ],
                                                    colony_location = initial_position, 
                                                    resource_layer = resource_layer) |>
        subset(Lon < maxx & Lon > minx) |>
        subset(Lat < maxy & Lat > miny) |>
        dplyr::slice_sample(n = 1) 
      if(is.null(nrow(potential_position))) { # if no points possible then do it again
        potential_position <- potential_position_func(n = 10, 
                                                      bearing = travel_bearing, 
                                                      step = travel_step, 
                                                      from = track_output[i - 1, ],
                                                      colony_location = initial_position, 
                                                      resource_layer = resource_layer) |>
          subset(Lon < maxx & Lon > minx) |>
          subset(Lat < maxy & Lat > miny) |>
          dplyr::slice_sample(n = 1) 
      } 
      track_output[i, c("Lon","Lat","angle","step","env","dist_col")] <- potential_position[, c("Lon","Lat","angle","step","env","dist_col")]
      track_output[i, "stepID"] <-  track_output[i-1, "stepID"] + 1
      track_output[i, "ymd_hms"] <-  track_output[i-1, "ymd_hms"] + lubridate::minutes(step_duration)
      track_output[i, "x_from"] <-  track_output[i-1, "Lon"]
      track_output[i, "y_from"] <-  track_output[i-1, "Lat"]
      track_output[i, "date_from"] <-  track_output[i-1, "ymd_hms"]
      track_output[i, "activity"] <- "travel"
    }else{
    ### Starts foraging bouts
      # sample potential positions based on movement type of previous step and select next step based on environment
      if(track_output[i-1, "activity"] == "travel"){
        potential_position <- potential_position_func(n = 10, 
                                                      bearing = travel_bearing, 
                                                      step = travel_step, 
                                                      from = track_output[i - 1, ],
                                                      colony_location = initial_position, 
                                                      resource_layer = resource_layer) |>
          dplyr::arrange(dplyr::desc(env)) |> # ordre d?croissant selon env
          dplyr::slice_head(n = 4) |> # prend les 4 premiers points
          subset(Lon < maxx & Lon > minx) |>
          subset(Lat < maxy & Lat > miny) 
          if(!is.null(nrow(potential_position))) {  # random select among the remaining points, if the table is not empty
            potential_position <- potential_position |> dplyr::slice_sample(n = 1) }
        if(is.null(nrow(potential_position))) { 
          potential_position <- potential_position_func(n = 10, 
                                                        bearing = travel_bearing, 
                                                        step = travel_step, 
                                                        from = track_output[i - 1, ],
                                                        colony_location = initial_position, 
                                                        resource_layer = resource_layer) |>
            dplyr::arrange(dplyr::desc(env)) |>
            dplyr::slice_head(n = 4) |>
            subset(Lon < maxx & Lon > minx) |>
            subset(Lat < maxy & Lat > miny)
            if(!is.null(nrow(potential_position))) {  
              potential_position <- potential_position |> dplyr::slice_sample(n = 1) }
        }
      }
      if(track_output[i-1, "activity"] == "forage"){
        potential_position <- potential_position_func(n = 10, 
                                                      bearing = foraging_bearing, 
                                                      step = foraging_step, 
                                                      from = track_output[i - 1, ],
                                                      colony_location = initial_position, 
                                                      resource_layer = resource_layer) |>
          dplyr::arrange(dplyr::desc(env)) |> # ordre d?croissant selon env
          dplyr::slice_head(n = 4) |> # prend les 4 premiers points
          subset(Lon < maxx & Lon > minx) |>
          subset(Lat < maxy & Lat > miny)
          if(!is.null(nrow(potential_position))) {  
            potential_position <- potential_position |> dplyr::slice_sample(n = 1) } #  subset(step == min(step))
        if(is.null(nrow(potential_position))) {  # if no points possible then do it again
          potential_position <- potential_position_func(n = 10, 
                                                        bearing = foraging_bearing, 
                                                        step = foraging_step, 
                                                        from = track_output[i - 1, ],
                                                        colony_location = initial_position, 
                                                        resource_layer = resource_layer) |>
            dplyr::arrange(dplyr::desc(env)) |>
            dplyr::slice_head(n = 4) |>
            subset(Lon < maxx & Lon > minx) |>
            subset(Lat < maxy & Lat > miny)
          if(!is.null(nrow(potential_position))) {  
            potential_position <- potential_position |> dplyr::slice_sample(n = 1) }
        }
      }
      track_output[i, c("Lon","Lat","angle","step","env","dist_col")] <- potential_position[, c("Lon","Lat","angle","step","env","dist_col")]
      track_output[i, "stepID"] <-  track_output[i-1, "stepID"] + 1
      track_output[i, "ymd_hms"] <-  track_output[i-1, "ymd_hms"] + lubridate::minutes(step_duration)
      track_output[i, "x_from"] <-  track_output[i-1, "Lon"]
      track_output[i, "y_from"] <-  track_output[i-1, "Lat"]
      track_output[i, "date_from"] <-  track_output[i-1, "ymd_hms"]
      # evaluate whether traveling or foraging (if env >= 0.5 = switch) for next step
      track_output[i, "activity"] <- ifelse( potential_position$env >= activity_threshold, "forage", "travel") 
    }
    i = i +1  
    }
  ### Starts homing phase
  i <- nrow(track_output) +1 
  while(track_output[i-1, "dist_col"] >= 0.5 ) {
    # sample potential positions based on movement type of previous step, select next step based on minimum distance to colony
    if(track_output[i-1, "activity"] == "travel"){
      potential_position <- potential_position_func(n = 10, 
                                                    bearing = travel_bearing, 
                                                    step = travel_step, 
                                                    from = track_output[i - 1, ],
                                                    colony_location = initial_position, 
                                                    resource_layer = resource_layer)  |>
        subset(Lon < maxx & Lon > minx) |>
        subset(Lat < maxy & Lat > miny) |>
        subset(dist_col == min(dist_col)) 
      if(nrow(potential_position > 1)) { potential_position <- potential_position[sample(1:nrow(potential_position), 1, replace = F),]  }
      if(is.null(nrow(potential_position))) { # if no points possible then do it again 
        potential_position <- potential_position_func(n = 10, 
                                                      bearing = travel_bearing, 
                                                      step = travel_step, 
                                                      from = track_output[i - 1, ],
                                                      colony_location = initial_position, 
                                                      resource_layer = resource_layer) |>
        subset(Lon < maxx & Lon > minx) |>
        subset(Lat < maxy & Lat > miny) |>
        subset(dist_col == min(dist_col))  
        if(nrow(potential_position > 1)) { potential_position <- potential_position[sample(1:nrow(potential_position), 1, replace = F),]  }
      }
    }
    if(track_output[i-1, "activity"] == "forage"){
      potential_position <- potential_position_func(n = 10, 
                                                    bearing = foraging_bearing, 
                                                    step = foraging_step, 
                                                    from = track_output[i - 1, ],
                                                    colony_location = initial_position, 
                                                    resource_layer = resource_layer)  |>
        subset(Lon < maxx & Lon > minx) |>
        subset(Lat < maxy & Lat > miny) |>
        subset(dist_col == min(dist_col))  
      if(nrow(potential_position > 1)) { potential_position <- potential_position[sample(1:nrow(potential_position), 1, replace = F),]  }
      if(is.null(nrow(potential_position))) {  # if no points possible then do it again
        potential_position <- potential_position_func(n = 10, 
                                                      bearing = foraging_bearing, 
                                                      step = foraging_step, 
                                                      from = track_output[i - 1, ],
                                                      colony_location = initial_position, 
                                                      resource_layer = resource_layer) |>
        subset(Lon < maxx & Lon > minx) |>
        subset(Lat < maxy & Lat > miny) |>
        subset(dist_col == min(dist_col))  
        if(nrow(potential_position > 1)) { potential_position <- potential_position[sample(1:nrow(potential_position), 1, replace = F),]  }
      }
    }
    track_output[i, c("Lon","Lat","angle","step","env","dist_col")] <- potential_position[, c("Lon","Lat","angle","step","env","dist_col")]
      track_output[i, "stepID"] <-  track_output[i-1, "stepID"] + 1
    track_output[i, "ymd_hms"] <-  track_output[i-1, "ymd_hms"] + lubridate::minutes(step_duration)
    track_output[i, "x_from"] <-  track_output[i-1, "Lon"]
    track_output[i, "y_from"] <-  track_output[i-1, "Lat"]
    track_output[i, "date_from"] <-  track_output[i-1, "ymd_hms"]
    track_output[i, "activity"] <- ifelse( potential_position$env >= activity_threshold, "forage", "travel")
    
    i <- i +1
  }
  # add point at colony to end the track
  track_output[i, c("Lon", "Lat")] <- initial_position[, c("Lon", "Lat")]
  track_output[i, "dist_col"] <-  0
  track_output[i, "stepID"] <-  track_output[i-1, "stepID"] + 1
  track_output[i, "ymd_hms"] <-  track_output[i-1, "ymd_hms"] + lubridate::minutes(step_duration)
  track_output[i, "x_from"] <-  track_output[i-1, "Lon"]
  track_output[i, "y_from"] <-  track_output[i-1, "Lat"]
  track_output[i, "date_from"] <-  track_output[i-1, "ymd_hms"]
  track_output[i, "activity"] <- "end"
  return(track_output)
}
# inherits checks for potential_positions
```
  
```{r example-simulate_trajectory_CPF}
library(ggplot2)
library(viridis)
library(tidyterra)
library(lubridate)
library(terra)
colony_location <- data.frame(Lon = 20, Lat = 20)
cdt <- generate_env_layer(grid = create_grid(), n = 1, seed = 25)
daylength <- insol::daylength(long = colony_location$Lon, 
                              lat = colony_location$Lat, 
                              jd = insol::JD(as.POSIXct("2022-08-01", tz = "UTC")), tmz = 0)
sunrise <- format(as.POSIXct(daylength[,1]*3600, 
                             origin = as.POSIXct("2022-08-01", tz = "UTC"), "%H:%M", tz = "UTC"))
    
# launch a travel for a duration of 12h
single_traj <- simulate_trajectory_CPF(initial_position = colony_location, 
                    resource_layer = cdt$rasters, 
                    starting_hour = ymd_hms(sunrise), # departs at sunrise
                    starting_bearing = c(90,10), 
                    starting_step = c(4.5, 3),
                    travel_bearing = c(0, 20), 
                    travel_step = c(3, 3), 
                    foraging_bearing = c(0, 0.5), 
                    foraging_step = c(1, 3),
                    minx = 0, maxx = 90, 
                    miny = 0, maxy = 90,
                    max_dist = 40,
                    step_duration = 5,
                    activity_threshold = 0.70,
                    max_duration = 720)

# view the trajectory
ggplot(single.traj) +
  geom_spatraster(data = cdt$rasters) + 
  geom_point(aes(x = Lon, y = Lat, color = activity)) +
  geom_point(data = colony_location, aes(x = Lon, y = Lat), col = "red") +
  scale_fill_viridis(option = "H") 

# look at the density distribution of distance to colony and movement parameters
ggplot(single_traj) + geom_density(aes(x = dist_col))
ggplot(single_traj |> subset(activity %in% c("forage", "travel"))) + geom_density(aes(x = angle)) + facet_wrap("activity")
ggplot(single_traj |> subset(activity %in% c("forage", "travel"))) + geom_density(aes(x = step)) + facet_wrap("activity")

```
  
```{r tests-simulate_trajectory_CPF}
test_that("simulate_trajectory works", {
  expect_true(inherits(simulate_trajectory_CPF, "function")) 
  
  library(ggplot2)
  library(viridis)
  library(tidyterra)
  library(lubridate)
  colony_location <- data.frame(Lon = 50, Lat = 50)
  cdt <- generate_env_layer(grid = create_grid(), n = 1, seed = 25)
  daylength <- insol::daylength(long = colony_location$Lon, 
                              lat = colony_location$Lat, 
                              jd = insol::JD(as.POSIXct("2022-08-01", tz = "UTC")), tmz = 0)
sunrise <- format(as.POSIXct(daylength[,1]*3600, 
                             origin = as.POSIXct("2022-08-01", tz = "UTC"), "%H:%M", tz = "UTC"))

  expect_true(
    is.data.frame(simulate_trajectory_CPF(initial_position = colony_location, 
                    resource_layer = cdt$rasters, 
                    starting_hour = ymd_hms(sunrise), 
                    starting_bearing = c(90,10), 
                    starting_step = c(4.5, 3),
                    travel_bearing = c(0, 20), 
                    travel_step = c(3, 3), 
                    foraging_bearing = c(0, 0.5), 
                    foraging_step = c(1, 3),
                    minx = 0, maxx = 90, 
                    miny = 0, maxy = 90,
                    max_dist = 40,
                    step_duration = 5,
                    activity_threshold = 0.70,
                    max_duration = 100)))
  
    expect_error(simulate_trajectory_CPF(initial_position = colony_location, 
                    resource_layer = colony_location, 
                    starting_hour = ymd_hms(sunrise), 
                    starting_bearing = c(90,10), 
                    starting_step = c(4.5, 3),
                    travel_bearing = c(0, 20), 
                    travel_step = c(3, 3), 
                    foraging_bearing = c(0, 0.5), 
                    foraging_step = c(1, 3),
                    minx = 0, maxx = 90, 
                    miny = 0, maxy = 90,
                    max_dist = 40,
                    step_duration = 5,
                    activity_threshold = 0.70,
                    max_duration = 100))
    
        expect_error(simulate_trajectory_CPF(initial_position = data.frame(Lon = c(50,60), Lat = c(50,60)), 
                        resource_layer = cdt$rasters, 
                        starting_hour = ymd_hms(sunrise), 
                        starting_bearing = c(90,10), 
                        starting_step = c(4.5, 3),
                        travel_bearing = c(0, 20), 
                        travel_step = c(3, 3), 
                        foraging_bearing = c(0, 0.5), 
                        foraging_step = c(1, 3),
                        minx = 0, maxx = 40, 
                        miny = 0, maxy = 40,
                        max_dist = 40,
                        step_duration = 5,
                        activity_threshold = 0.70,
                        max_duration = 100))
})
```
  

# Simulate the trajectory of a non-central place forager

`simulate_trajectory_FR` does the same as above, but the individual is not restrained around its colony and does not perform homing bout. Should add a condition on the time spent aorund a same area. 
    
```{r function-simulate_trajectory_FR}
#' Simulate individual trajectory
#'
#' @param initial_position data.frame containing the coordinates of the point the individual starts from (Lon, Lat fields)
#' @param resource_layer SpatRaster of the resource layer
#' @param starting_hour Starting hour
#' @param starting_bearing Bearing when departing the colony (mu, kappa)
#' @param starting_step Step length when departing the colony (mu, kappa)
#' @param travel_bearing Bearing from previous step when traveling (mu, kappa)
#' @param travel_step Distance from previous step when traveling (scale, rate) 
#' @param foraging_bearing Bearing from previous step when foraging/searching (mu, kappa)
#' @param foraging_step Distance from previous step when foraging (scale, rate)
#' @param minx Minimum longitude of the bounding box to simulate trajectory within
#' @param maxx Maximum longitude of the bounding box to simulate trajectory within
#' @param miny Minimum latitude of the bounding box to simulate trajectory within
#' @param maxy Maximum latitude of the bounding box to simulate trajectory within
#' @param max_duration Total duration of a full trip, in minutes (defaults to 720, 12h)
#' @param step_duration Duration of a step length, in minutes (defaults 1), defines the sampling interval
#' @param activity_threshold Threshold environmental value defining when an individual switches from travelling to foraging movement
#'
#' @import lubridate
#' @importFrom dplyr full_join slice_sample slice_head desc arrange
#' 
#' @return a data.frame
#' @export
#'
#' @examples
simulate_trajectory_FR <- function(initial_position, 
                                resource_layer, 
                       starting_hour,
                       starting_bearing = c(90,10),
                       starting_step = c(4.5, 3),
                       travel_bearing = c(0, 20), # (mu, kappa) -> higher kappa = higher concentration
                       travel_step = c(3, 3), #  (scale, rate) -> smaller the rate, higher the dispersion
                       foraging_bearing = c(0, 0.5), 
                       foraging_step = c(1, 3), 
                       minx = 0, maxx = 90, 
                       miny = 0, maxy = 90,
                       step_duration = 1,
                       activity_threshold = 0.5,
                       max_duration = 720) 
{
    if(isFALSE(initial_position$Lon >= minx & initial_position$Lon <= maxx &
         initial_position$Lat >= miny & initial_position$Lat <= maxy)){
    stop("initial_position must be included in the bounding box given by minx, maxx, miny and maxy")
  }
  ext_coord <- as.vector(terra::ext(resource_layer))
  if(isFALSE(initial_position$Lon >= ext_coord[[1]] & initial_position$Lon <= ext_coord[[2]] &
            initial_position$Lat >= ext_coord[[3]] & initial_position$Lat <= ext_coord[[4]])){
    stop("initial_position must be included in the resource_layer extent")
  }
  if(nrow(initial_position) != 1){
    stop("initial_position must include a single row")
  }
  
  track_output <- data.frame(Lon = initial_position$Lon,
                             Lat = initial_position$Lat,
                             angle = NA,
                             step = 0,
                             env = NA,
                             dist_col = 0,
                             activity = "start",
                             stepID = 0,
                             ymd_hms = starting_hour,
                             x_from = initial_position$Lon,
                             y_from = initial_position$Lat,
                             date_from = starting_hour)
  
  ### Departs from the colony - outbound phase of travel and foraging bouts
  i <- 2
  while(track_output[i-1, "ymd_hms"] < ymd_hms(starting_hour) + lubridate::minutes(max_duration)
  ){
    # starts with the initial bearing
    if(i == 2){ 
      potential_position <- potential_position_func(n = 10, 
                                                    bearing = starting_bearing, 
                                                    step = starting_step, 
                                                    from = track_output[i - 1, ],
                                                    colony_location = initial_position, 
                                                    resource_layer = resource_layer) |>
        subset(Lon < maxx & Lon > minx) |>
        subset(Lat < maxy & Lat > miny) |>
        dplyr::slice_sample(n = 1)
      if(is.null(nrow(potential_position))) { # if no points possible then do it again
        potential_position <- potential_position_func(n = 10, 
                                                      bearing = starting_bearing, 
                                                      step = starting_step, 
                                                      from = track_output[i - 1, ],
                                                      colony_location = initial_position, 
                                                      resource_layer = resource_layer) |>
          subset(Lon < maxx & Lon > minx) |>
          subset(Lat < maxy & Lat > miny) |>
          dplyr::slice_sample(n = 1)
        } 
      track_output[i, c("Lon","Lat","angle","step","env","dist_col")] <- potential_position[, c("Lon","Lat","angle","step","env","dist_col")]
      track_output[i, "stepID"] <-  track_output[i-1, "stepID"] + 1
      track_output[i, "ymd_hms"] <-  track_output[i-1, "ymd_hms"] + lubridate::minutes(step_duration)
      track_output[i, "x_from"] <-  track_output[i-1, "Lon"]
      track_output[i, "y_from"] <-  track_output[i-1, "Lat"]
      track_output[i, "date_from"] <-  track_output[i-1, "ymd_hms"]
      track_output[i, "activity"] <- "travel"
    }
    # continue with traveling parameters but not yet condition on environment
    if(i %in% c(3:5)){
      potential_position <- potential_position_func(n = 10, 
                                                    bearing = travel_bearing, 
                                                    step = travel_step, 
                                                    from = track_output[i - 1, ],
                                                    colony_location = initial_position, 
                                                    resource_layer = resource_layer) |>
        subset(Lon < maxx & Lon > minx) |>
        subset(Lat < maxy & Lat > miny) |>
        dplyr::slice_sample(n = 1) 
      if(is.null(nrow(potential_position))) { # if no points possible then do it again
        potential_position <- potential_position_func(n = 10, 
                                                      bearing = travel_bearing, 
                                                      step = travel_step, 
                                                      from = track_output[i - 1, ],
                                                      colony_location = initial_position, 
                                                      resource_layer = resource_layer) |>
          subset(Lon < maxx & Lon > minx) |>
          subset(Lat < maxy & Lat > miny) |>
          dplyr::slice_sample(n = 1) 
      } 
      track_output[i, c("Lon","Lat","angle","step","env","dist_col")] <- potential_position[, c("Lon","Lat","angle","step","env","dist_col")]
      track_output[i, "stepID"] <-  track_output[i-1, "stepID"] + 1
      track_output[i, "ymd_hms"] <-  track_output[i-1, "ymd_hms"] + lubridate::minutes(step_duration)
      track_output[i, "x_from"] <-  track_output[i-1, "Lon"]
      track_output[i, "y_from"] <-  track_output[i-1, "Lat"]
      track_output[i, "date_from"] <-  track_output[i-1, "ymd_hms"]
      track_output[i, "activity"] <- "travel"
    }else{
    ### Starts foraging bouts
      # sample potential positions based on movement type of previous step and select next step based on environment
      if(track_output[i-1, "activity"] == "travel"){
        potential_position <- potential_position_func(n = 10, 
                                                      bearing = travel_bearing, 
                                                      step = travel_step, 
                                                      from = track_output[i - 1, ],
                                                      colony_location = initial_position, 
                                                      resource_layer = resource_layer) |>
          dplyr::arrange(dplyr::desc(env)) |> # ordre d?croissant selon env
          dplyr::slice_head(n = 4) |> # prend les 4 premiers points
          subset(Lon < maxx & Lon > minx) |>
          subset(Lat < maxy & Lat > miny) 
          if(!is.null(nrow(potential_position))) {  # random select among the remaining points, if the table is not empty
            potential_position <- potential_position |> dplyr::slice_sample(n = 1) }
        if(is.null(nrow(potential_position))) { 
          potential_position <- potential_position_func(n = 10, 
                                                        bearing = travel_bearing, 
                                                        step = travel_step, 
                                                        from = track_output[i - 1, ],
                                                        colony_location = initial_position, 
                                                        resource_layer = resource_layer) |>
            dplyr::arrange(dplyr::desc(env)) |>
            dplyr::slice_head(n = 4) |>
            subset(Lon < maxx & Lon > minx) |>
            subset(Lat < maxy & Lat > miny)
            if(!is.null(nrow(potential_position))) {  
              potential_position <- potential_position |> dplyr::slice_sample(n = 1) }
        }
      }
      if(track_output[i-1, "activity"] == "forage"){
        potential_position <- potential_position_func(n = 10, 
                                                      bearing = foraging_bearing, 
                                                      step = foraging_step, 
                                                      from = track_output[i - 1, ],
                                                      colony_location = initial_position, 
                                                      resource_layer = resource_layer) |>
          dplyr::arrange(dplyr::desc(env)) |> # ordre d?croissant selon env
          dplyr::slice_head(n = 4) |> # prend les 4 premiers points
          subset(Lon < maxx & Lon > minx) |>
          subset(Lat < maxy & Lat > miny)
          if(!is.null(nrow(potential_position))) {  
            potential_position <- potential_position |> dplyr::slice_sample(n = 1) } #  subset(step == min(step))
        if(is.null(nrow(potential_position))) {  # if no points possible then do it again
          potential_position <- potential_position_func(n = 10, 
                                                        bearing = foraging_bearing, 
                                                        step = foraging_step, 
                                                        from = track_output[i - 1, ],
                                                        colony_location = initial_position, 
                                                        resource_layer = resource_layer) |>
            dplyr::arrange(dplyr::desc(env)) |>
            dplyr::slice_head(n = 4) |>
            subset(Lon < maxx & Lon > minx) |>
            subset(Lat < maxy & Lat > miny)
          if(!is.null(nrow(potential_position))) {  
            potential_position <- potential_position |> dplyr::slice_sample(n = 1) }
        }
      }
      track_output[i, c("Lon","Lat","angle","step","env","dist_col")] <- potential_position[, c("Lon","Lat","angle","step","env","dist_col")]
      track_output[i, "stepID"] <-  track_output[i-1, "stepID"] + 1
      track_output[i, "ymd_hms"] <-  track_output[i-1, "ymd_hms"] + lubridate::minutes(step_duration)
      track_output[i, "x_from"] <-  track_output[i-1, "Lon"]
      track_output[i, "y_from"] <-  track_output[i-1, "Lat"]
      track_output[i, "date_from"] <-  track_output[i-1, "ymd_hms"]
      # evaluate whether traveling or foraging (if env >= 0.5 = switch) for next step
      track_output[i, "activity"] <- ifelse( potential_position$env >= activity_threshold, "forage", "travel") 
    }
    i = i +1  
    }
  return(track_output)
}
# inherits checks for potential.positions
```
  
```{r example-simulate_trajectory_FR}
library(ggplot2)
library(viridis)
library(tidyterra)
library(lubridate)
colony_location <- data.frame(Lon = 50, Lat = 50)
cdt <- generate_env_layer(grid = create_grid(), n = 1, seed = 4)
terra::plot(cdt$rasters)
# determine the sunrise hour for a random dat
daylength <- insol::daylength(long = colony_location$Lon, 
                              lat = colony_location$Lat, 
                              jd = insol::JD(as.POSIXct("2022-08-01", tz = "UTC")), tmz = 0)
sunrise <- format(as.POSIXct(daylength[,1]*3600, 
                             origin = as.POSIXct("2022-08-01", tz = "UTC"), "%H:%M", tz = "UTC"))
    
single_traj <- simulate_trajectory_FR(initial_position = colony_location, 
                    resource_layer = cdt$rasters, 
                    starting_hour = ymd_hms(sunrise), # departs at sunrise
                    starting_bearing = c(90,10), 
                    starting_step = c(4.5, 3),
                    travel_bearing = c(0, 20), 
                    travel_step = c(3, 3), 
                    foraging_bearing = c(0, 0.5), 
                    foraging_step = c(1, 3),
                    minx = 0, maxx = 90, 
                    miny = 0, maxy = 90,
                    step_duration = 1,
                    activity_threshold = 0.7,
                    max_duration = 2000)

# view the trajectory
ggplot(single_traj) +
  geom_spatraster(data = cdt$rasters) + 
  geom_point(aes(x = Lon, y = Lat, color = activity)) +
  geom_point(data = colony_location, aes(x = Lon, y = Lat), col = "red") +
  scale_fill_viridis(option = "H") 

# look at the density distribution of distance to colony and movement parameters
ggplot(single_traj) + geom_density(aes(x = dist_col))
ggplot(single_traj |> subset(activity %in% c("forage", "travel"))) + geom_density(aes(x = angle)) + facet_wrap("activity")
ggplot(single_traj |> subset(activity %in% c("forage", "travel"))) + geom_density(aes(x = step)) + facet_wrap("activity")


```
  
```{r tests-simulate_trajectory_FR}
test_that("simulate_trajectory works", {
  expect_true(inherits(simulate_trajectory_CPF, "function")) 
  
  library(ggplot2)
  library(viridis)
  library(tidyterra)
  library(lubridate)
  colony_location <- data.frame(Lon = 50, Lat = 50)
  cdt <- generate_env_layer(grid = create_grid(), n = 1, seed = 25)
  daylength <- insol::daylength(long = colony_location$Lon, 
                              lat = colony_location$Lat, 
                              jd = insol::JD(as.POSIXct("2022-08-01", tz = "UTC")), tmz = 0)
sunrise <- format(as.POSIXct(daylength[,1]*3600, 
                             origin = as.POSIXct("2022-08-01", tz = "UTC"), "%H:%M", tz = "UTC"))

  expect_true(
    is.data.frame(simulate_trajectory_FR(initial_position = colony_location, 
                    resource_layer = cdt$rasters, 
                    starting_hour = ymd_hms(sunrise), # departs at sunrise
                    starting_bearing = c(90,10), 
                    starting_step = c(4.5, 3),
                    travel_bearing = c(0, 20), 
                    travel_step = c(3, 3), 
                    foraging_bearing = c(0, 0.5), 
                    foraging_step = c(1, 3),
                    minx = 0, maxx = 90, 
                    miny = 0, maxy = 90,
                    step_duration = 1,
                    activity_threshold = 0.7,
                    max_duration = 100)))
  
  expect_error(simulate_trajectory_FR(initial_position = colony_location, 
                    resource_layer = colony_location, 
                    starting_hour = ymd_hms(sunrise), 
                    starting_bearing = c(90,10), 
                    starting_step = c(4.5, 3),
                    travel_bearing = c(0, 20), 
                    travel_step = c(3, 3), 
                    foraging_bearing = c(0, 0.5), 
                    foraging_step = c(1, 3),
                    minx = 0, maxx = 90, 
                    miny = 0, maxy = 90,
                    step_duration = 5,
                    activity_threshold = 0.70,
                    max_duration = 100))
  
  expect_error(simulate_trajectory_FR(initial_position = data.frame(Lon = c(50,60), Lat = c(50,60)), 
                        resource_layer = cdt$rasters, 
                        starting_hour = ymd_hms(sunrise), 
                        starting_bearing = c(90,10), 
                        starting_step = c(4.5, 3),
                        travel_bearing = c(0, 20), 
                        travel_step = c(3, 3), 
                        foraging_bearing = c(0, 0.5), 
                        foraging_step = c(1, 3),
                        minx = 0, maxx = 40, 
                        miny = 0, maxy = 40,
                        step_duration = 5,
                        activity_threshold = 0.70,
                        max_duration = 100))
  

})
```
  



```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_simulate_mvmt.Rmd", vignette_name = "Simulate individual movements")
```

